Why aren't significance tests commonly used for linear regression diagnostics?

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.Rmd
regroup sections
update p value calc

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \\ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

@@

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

Data used in the $m - 1$ null plots needs to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

```{r first-example-lineup, fig.cap="Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?"}
mod <- poly_model(include_z = TRUE, sigma = 1)
POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

@@

### Lineup protocol {#se:lineup}

With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error. Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test. When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature. It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed. Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009].

Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level. However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided.

@buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$. It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup.

Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$. If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.

This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen. A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009]. Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians.

The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$. In fact, the lower bound of $p$-value decreases exponentially as $K$ increases. With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$. Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected.

As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$.

The full model was a Dirichlet-multinomial mixture distribution

```{=tex}
\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
\end{split}\end{align}
```
Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution

```{=tex}
\begin{align} \label{eq:beta-binomial}\begin{split}
\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
C_i&|\theta_i \sim Binomial(K, \theta_i).
\end{split}\end{align}
```
Thus, the visual p-value followed by the beta-binomial model is given as

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}
```
where $B(.)$ is the beta function defined as

```{=tex}
\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}
```
Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment.

Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package. We assume the attractiveness of the plot $i$ modelled as $w_i \sim Uniform(0,1)$ for $i=1,..,m$. Let $\theta_i = w_i/\sum_{i=1}^{m}w_i$ be the probability of plot $i$ being selected by a subject. Then, given the number of selections $s_j$, for $j=1,...,K$, the distribution of $C_i$ can be approximated by simulating the random selection process with computer. The simulated visual test p-value is formulated as

```{=tex}
\begin{equation} \label{eq:p-value-multi}
\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
\end{equation}
```
<!-- ### Effectiveness of visual inference in regression diagnostics -->

>>>>

Graphical summaries in which residuals are plotted against fitted values or other functions of the predictor variables that are approximately orthogonal to residuals are referred to as standard residual plots in @cook1982residuals. As shown in Figure \ref{fig:commondepartures}, the top-left panel is a good residual plot with residuals evenly distributed at both sides of the horizontal zero line showing no noticeable patterns. There are various types of departures from a good residual plot. We will only discuss three most commonly checked departures in this paper. Namely, non-linearity, heterskedasticity and non-normality.

@@

### Lineup protocol

@@

Visual inference was first introduced in a 1999 Joint Statistical Meetings (JSM) talk with the title "Inference for Data Visualization" by @buja_inference_1999 as an idea to address the issue of valid inference for visual discoveries of data plots [@gelman_exploratory_2004]. Later, @buja_statistical_2009 proposed the lineup protocol as a visual test inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$. If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$.

```{r first-example-lineup, fig.cap="An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?"}
mod <- poly_model(include_z = TRUE, sigma = 1)
POLY_MODEL$plot_lineup(mod$gen_lineup(300, pos = 11), theme = theme_light(), remove_axis = TRUE, remove_grid_line = TRUE)
```

Figure \ref{fig:first-example-lineup} is an example of a lineup protocol. If the actual data plot at position 11 is identifiable, then it is evidence for the rejection of the null hypothesis that the regression model is correctly specified. In fact, the actual residual plot is obtained from a misspecified regression model with non-linearity issue.

### Sampling from the null distribution

Data used in the $m - 1$ null plots need to be simulated. In the context of regression diagnostics, sampling data from $H_0$ is equivalent to sampling data from the assumed model. As @buja_statistical_2009 suggested, $H_0$ is usually composited by a collection of distributions controlled by nuisance parameters. Since regression models can have various forms, there is no general solution to this problem, but it sometimes can be reduced to so called "reference distribution" by applying one of the three methods: (i) sampling from a conditional distribution given a minimal sufficient statistic under $H_0$, (ii) parametric bootstrap sampling with nuisance parameters estimated under $H_0$, and (iii) Bayesian posterior predictive sampling. The conditional distribution given a minimal sufficient statistic is the best justified reference distribution among the three [@buja_statistical_2009]. Essentially, null residuals can be simulated by regressing $N$ i.i.d standard normal random draws on the regressors, then rescaling it by the ratio of residual sum of square in two regressions.

### p-value calculation

As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. Thus, the visual p-value followed by the above model is given as

```{=tex}
\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha),
\end{equation}
```

where $B(.)$ is the beta function defined as

```{=tex}
\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0. 
\end{equation}
```

@@

<!-- ### Lineup protocol {#se:lineup} -->

<!-- With the simulation mechanism of null plots being provided, another aspect of hypothesis testing that needs to be addressed is the control of false positive rate or Type I error. Any visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ needs to pair with a critical value $c^{(i)}$ to form a hypothesis test. When a visual feature $i$ is discovered by the observer from a plot, the corresponding visual statistic $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known as there is no general agreement on the measurement of the degree of presence of a visual feature. It is only the event that $T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ is confirmed. Similarly, if any visual discovery is found by the observer, we say, there exists $i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}$ [@buja_statistical_2009]. -->

<!-- Using the above definition, the family-wise Type I error can be controlled if one can provide the collection of critical values $c^{(i)}~(i \in I)$ such that $P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha$, where $\alpha$ is the significance level. However, since the quantity of $T^{(i)}(\boldsymbol{\mathrm{y}})$ may not be known, such collection of critical values can not be provided. -->

<!-- @buja_statistical_2009 proposed the lineup protocol as a visual test to calibrate the Type I error issue without the specification of $c^{(i)}~(i \in I)$. It is inspired by the "police lineup" or "identity parade" which is the act of asking the eyewitness to identify criminal suspect from a group of irrelevant people. The protocol consists of $m$ randomly placed plots, where one plot is the actual data plot, and the remaining $m - 1$ plots have the identical graphical production as the data plot except the data has been replaced with data consistent with the null hypothesis. Then, an observer who have not seen the actual data plot will be asked to point out the most different plot from the lineup. -->

<!-- Under the null hypothesis, it is expected that the actual data plot would have no distinguishable difference with the null plots, and the probability of the observer correctly picks the actual data plot is $1/m$. If we reject the null hypothesis as the observer correctly picks the actual data plot, then the Type I error of this test is $1/m$. -->

<!-- This provides us with an mechanism to control the Type I error, because $m$ - the number of plots in a lineup can be chosen. A larger value of $m$ will result in a smaller Type I error, but the limit to the value of $m$ depends on the number of plots a human is willing to view [@buja_statistical_2009]. Typically, $m$ will be set to $20$ which is equivalent to set $\alpha = 0.05$, a general choice of significance level for conventional testing among statisticians. -->

<!-- Further, a visual test can involve $K$ independent observers. Let $D_i = \{0,1\}$ be a binomial random variable denoting whether subject $i$ correctly detecting the actual data plot, and $X = \sum_{i=1}^{K}X_i$ be the number of observers correctly picking the actual data plot. Then, by imposing a relatively strong assumption on the visual test that all $K$ evaluations are fully independent, under the null hypothesis, $X \sim \mathrm{Binom}_{K,1/m}$. Therefore, the $p$-value of a lineup of size $m$ evaluated by $K$ observer is given as \begin{equation} \label{eq:pvaluesingle} -->
<!-- P(X \geq x) = \sum_{i=x}^{K}{{K}\choose{i}}\left(\frac{1}{m}\right)^i\left(\frac{m-1}{m}\right)^{k-i}, -->
<!-- \end{equation} -->

<!-- where $x$ is the realization of number of observers correctly picking the actual data plot [@majumder_validation_2013]. -->

<!-- The multiple individuals approach avoids the limit of $m$, while provides visual tests with $p$-value much smaller than $0.05$. In fact, the lower bound of $p$-value decreases exponentially as $K$ increases. With just $4$ individuals and $20$ data plots in a lineup, the $p$-value could be as small as $0.0001$. Additionally, by involving multiple observers, variation of individual ability to read plots can be addressed to some degree as different opinions about visual discoveries can be collected. -->

<!-- As pointed out by @vanderplas2021statistical, though Equation (\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the possible dependencies in the visual test due to repeated evaluations of the same lineup. And it is inapplicable to visual test where subjects are asked to select one or more "most different" plots from the lineup. They summarized three common different scenarios in visual inference: (1) $K$ different lineups are shown to $K$ subjects, (2) $K$ lineups with different null plots but the same actual data plot are shown to $K$ subjects, and (3) the same lineup is shown to $K$ subjects. Out of these three scenarios, Scenario 3 is the most common in previous studies as it puts the least constraints on the experimental design. For Scenario 3, @vanderplas2021statistical modelled the probability of a plot $i$ being selected from a lineup as $\theta_i$, where $\theta_i \sim Dirichlet(\alpha)$ for $i=1,...,m$ and $\alpha > 0$. And defined $c_i$ to be the number of times plot $i$ being selected in $K$ evaluations. In case subject $j$ makes multiple selections, they decided to add $1/s_j$ to $c_i$ instead of one, where $s_j$ is the number of plots subject $j$ selected for $j=1,...K$. This ensured $\sum_{i}c_i=K$. -->

<!-- The full model was a Dirichlet-multinomial mixture distribution -->

<!-- ```{=tex} -->
<!-- \begin{align} \label{eq:dirichlet-multinomial}\begin{split} -->
<!-- \boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\ -->
<!-- (c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}). -->
<!-- \end{split}\end{align} -->
<!-- ``` -->
<!-- Since the p-value calculation only needs to concern the number of times the actual data plot being selected denoted by $C_i$, they showed the model can be simplified to a beta-binomial mixture distribution -->

<!-- ```{=tex} -->
<!-- \begin{align} \label{eq:beta-binomial}\begin{split} -->
<!-- \theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\ -->
<!-- C_i&|\theta_i \sim Binomial(K, \theta_i). -->
<!-- \end{split}\end{align} -->
<!-- ``` -->
<!-- Thus, the visual p-value followed by the beta-binomial model is given as -->

<!-- ```{=tex} -->
<!-- \begin{equation} \label{eq:pvalue-beta-binomial} -->
<!-- P(C \geq c_i) = \sum_{x=c_i}^{K}{{K}\choose{x}}\frac{1}{B(\alpha, (m-1)\alpha)}B(x + \alpha, K - x + (m - 1)\alpha), -->
<!-- \end{equation} -->
<!-- ``` -->
<!-- where $B(.)$ is the beta function defined as -->

<!-- ```{=tex} -->
<!-- \begin{equation} \label{eq:betafunction} -->
<!-- B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.  -->
<!-- \end{equation} -->
<!-- ``` -->

<!-- Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires the estimation of $\hat{\alpha}$, which largely depends on the null model, the type of the plot and other aesthetic features. They suggested to estimate $\hat{\alpha}$ visually based on the selections of null plots of the experimental data, or to estimate $\hat{\alpha}$ numerically based on several additional Rorschach lineups, which is a type of lineup containing only null plots. However, when the number of null models are large, it could be expensive to manually estimate each $\alpha$ or include additional Rorschach lineups in the experiment. -->

<!-- Instead, in the experiments that will be described in section \ref{experimental-design}, we adopt a simpler model implicitly used by the `pmulti()` function of the `vinference` `R` package. We assume the attractiveness of the plot $i$ modelled as $w_i \sim Uniform(0,1)$ for $i=1,..,m$. Let $\theta_i = w_i/\sum_{i=1}^{m}w_i$ be the probability of plot $i$ being selected by a subject. Then, given the number of selections $s_j$, for $j=1,...,K$, the distribution of $C_i$ can be approximated by simulating the random selection process with computer. The simulated visual test p-value is formulated as -->

<!-- ```{=tex} -->
<!-- \begin{equation} \label{eq:p-value-multi} -->
<!-- \text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}. -->
<!-- \end{equation} -->
<!-- ``` -->
<!-- ### Effectiveness of visual inference in regression diagnostics -->

### 2022-09-13 paper_comparison/paper_comparison/paper_comparison.tex
update p value calc

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\textbackslash ref\{fig:commondepartures\}, the top-left panel is a good
residual plot with residuals evenly distributed at both sides of the
horizontal zero line showing no noticeable patterns. There are various
types of departures from a good residual plot. We will only discuss
three most commonly checked departures in this paper. Namely,
non-linearity, heterskedasticity and non-normality.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{Example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?}\label{fig:first-example-lineup}
\end{figure}

@@

\hypertarget{se:lineup}{%
\subsubsection{Lineup protocol}\label{se:lineup}}

With the simulation mechanism of null plots being provided, another
aspect of hypothesis testing that needs to be addressed is the control
of false positive rate or Type I error. Any visual statistic
\(T^{(i)}(\boldsymbol{\mathrm{y}})\) needs to pair with a critical value
\(c^{(i)}\) to form a hypothesis test. When a visual feature \(i\) is
discovered by the observer from a plot, the corresponding visual
statistic \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known as there
is no general agreement on the measurement of the degree of presence of
a visual feature. It is only the event that
\(T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\) is confirmed. Similarly,
if any visual discovery is found by the observer, we say, there exists
\(i \in I:~T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}\)
\citep{buja_statistical_2009}.

Using the above definition, the family-wise Type I error can be
controlled if one can provide the collection of critical values
\(c^{(i)}~(i \in I)\) such that
\(P(\mathrm{there~exists~} i \in I: T^{(i)}(\boldsymbol{\mathrm{y}}) > c^{(i)}|\boldsymbol{\mathrm{y}}) \leq \alpha\),
where \(\alpha\) is the significance level. However, since the quantity
of \(T^{(i)}(\boldsymbol{\mathrm{y}})\) may not be known, such
collection of critical values can not be provided.

\citet{buja_statistical_2009} proposed the lineup protocol as a visual
test to calibrate the Type I error issue without the specification of
\(c^{(i)}~(i \in I)\). It is inspired by the ``police lineup'' or
``identity parade'' which is the act of asking the eyewitness to
identify criminal suspect from a group of irrelevant people. The
protocol consists of \(m\) randomly placed plots, where one plot is the
actual data plot, and the remaining \(m - 1\) plots have the identical
graphical production as the data plot except the data has been replaced
with data consistent with the null hypothesis. Then, an observer who
have not seen the actual data plot will be asked to point out the most
different plot from the lineup.

Under the null hypothesis, it is expected that the actual data plot
would have no distinguishable difference with the null plots, and the
probability of the observer correctly picks the actual data plot is
\(1/m\). If we reject the null hypothesis as the observer correctly
picks the actual data plot, then the Type I error of this test is
\(1/m\).

This provides us with an mechanism to control the Type I error, because
\(m\) - the number of plots in a lineup can be chosen. A larger value of
\(m\) will result in a smaller Type I error, but the limit to the value
of \(m\) depends on the number of plots a human is willing to view
\citep{buja_statistical_2009}. Typically, \(m\) will be set to \(20\)
which is equivalent to set \(\alpha = 0.05\), a general choice of
significance level for conventional testing among statisticians.

The multiple individuals approach avoids the limit of \(m\), while
provides visual tests with \(p\)-value much smaller than \(0.05\). In
fact, the lower bound of \(p\)-value decreases exponentially as \(K\)
increases. With just \(4\) individuals and \(20\) data plots in a
lineup, the \(p\)-value could be as small as \(0.0001\). Additionally,
by involving multiple observers, variation of individual ability to read
plots can be addressed to some degree as different opinions about visual
discoveries can be collected.

As pointed out by \citet{vanderplas2021statistical}, though Equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\).

The full model was a Dirichlet-multinomial mixture distribution

\begin{align} \label{eq:dirichlet-multinomial}\begin{split}
\boldsymbol{\theta}&|\alpha \sim Dirichlet(\alpha)\\
(c_i,...,c_m)&|\boldsymbol{\theta} \sim Multinomial(K, \boldsymbol{\theta}).
\end{split}\end{align}

Since the p-value calculation only needs to concern the number of times
the actual data plot being selected denoted by \(C_i\), they showed the
model can be simplified to a beta-binomial mixture distribution

\begin{align} \label{eq:beta-binomial}\begin{split}
\theta_i&|\alpha \sim Beta(\alpha, (m-1)\alpha)\\
C_i&|\theta_i \sim Binomial(K, \theta_i).
\end{split}\end{align}

Thus, the visual p-value followed by the beta-binomial model is given as

@@

Note that the use of Equation (\ref{eq:pvalue-beta-binomial}) requires
the estimation of \(\hat{\alpha}\), which largely depends on the null
model, the type of the plot and other aesthetic features. They suggested
to estimate \(\hat{\alpha}\) visually based on the selections of null
plots of the experimental data, or to estimate \(\hat{\alpha}\)
numerically based on several additional Rorschach lineups, which is a
type of lineup containing only null plots. However, when the number of
null models are large, it could be expensive to manually estimate each
\(\alpha\) or include additional Rorschach lineups in the experiment.

Instead, in the experiments that will be described in section
\ref{experimental-design}, we adopt a simpler model implicitly used by
the \texttt{pmulti()} function of the \texttt{vinference} \texttt{R}
package. We assume the attractiveness of the plot \(i\) modelled as
\(w_i \sim Uniform(0,1)\) for \(i=1,..,m\). Let
\(\theta_i = w_i/\sum_{i=1}^{m}w_i\) be the probability of plot \(i\)
being selected by a subject. Then, given the number of selections
\(s_j\), for \(j=1,...,K\), the distribution of \(C_i\) can be
approximated by simulating the random selection process with computer.
The simulated visual test p-value is formulated as

\begin{equation} \label{eq:p-value-multi}
\text{p-value} = \frac{\#draws~that~the~actual~data~plota~being~selected~more~than~c_i~times}{\#simulation}.
\end{equation}

>>>>

Graphical summaries in which residuals are plotted against fitted values
or other functions of the predictor variables that are approximately
orthogonal to residuals are referred to as standard residual plots in
\citet{cook1982residuals}. As shown in Figure
\ref{fig:commondepartures}, the top-left panel is a good residual plot
with residuals evenly distributed at both sides of the horizontal zero
line showing no noticeable patterns. There are various types of
departures from a good residual plot. We will only discuss three most
commonly checked departures in this paper. Namely, non-linearity,
heterskedasticity and non-normality.

@@

\begin{figure}
\includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/first-example-lineup-1} \caption{An example lineup consists of 20 randomly placed residual plots, where 19 are null plots with data consistent with the null hypothesis, and one is the actual residual plot. Can you find the most different plot?}\label{fig:first-example-lineup}
\end{figure}

@@

As pointed out by \citet{vanderplas2021statistical}, though Equation
(\ref{eq:pvaluesingle}) is trivial, but it doesn't take into account the
possible dependencies in the visual test due to repeated evaluations of
the same lineup. And it is inapplicable to visual test where subjects
are asked to select one or more ``most different'' plots from the
lineup. They summarized three common different scenarios in visual
inference: (1) \(K\) different lineups are shown to \(K\) subjects, (2)
\(K\) lineups with different null plots but the same actual data plot
are shown to \(K\) subjects, and (3) the same lineup is shown to \(K\)
subjects. Out of these three scenarios, Scenario 3 is the most common in
previous studies as it puts the least constraints on the experimental
design. For Scenario 3, \citet{vanderplas2021statistical} modelled the
probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). And defined \(c_i\) to be the number
of times plot \(i\) being selected in \(K\) evaluations. In case subject
\(j\) makes multiple selections, they decided to add \(1/s_j\) to
\(c_i\) instead of one, where \(s_j\) is the number of plots subject
\(j\) selected for \(j=1,...K\). This ensured \(\sum_{i}c_i=K\). Thus,
the visual p-value followed by the beta-binomial model is given as

@@
